Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})

### 2019-10-03 manuscript/pest-management-science/main.tex
more author feedback...

Onion maggot (\textit{Delia antiqua}) is a pest of onions worldwide. Current means of managing this pest rely heavily on prophylatic insecticide treatments at planting. These options may not be viable in organic production systems or areas where insecticide-resistant populations occur. Here we explore the efficacy of an attract and kill strategy for control of \textit{D. antiqua} evaluating the ability of attractive, spinosad containing spheres to kill adult \textit{D. antiqua} and reduce crop losses. Spinosad containing spheres were able to consistently kill \textit{D. antiqua} adults over the course of the field season (mortality range: between 49\% and 59\% on average). Pairing spinosad spheres with Delia Lure increased efficacy by 72\% compared with the attract and kill strategy alone. Performance of this attract and kill strategy also can reduce damage by \textit{D. antiqua}  larvae in the field, but it did not achieve a level of control comparable to the level provided by a conventional insecticide treatment. Implementation of this attract and kill strategy could be a valuable tool in situations where conventional pesticides are either not available or desired, where additional control techniques are needed, or to provide a season-long option for control of \textit{D. antiqua} populations.

@@

The lifecycle of \textit{Delia antiqua} plays a large role in its severity as an onion pest.  In the northern United States, three generations per year are common \citep{eckenrode1975population, hoepting2004insecticide}. In New York, \textit{D. antiqua} overwinters as a pupa and adults emerge in late April to early May and peak in late May to early June. The next generation begins to emerge in late June to early July and peaks in mid to late July, while the last generation adults begin to emerge in late July to early August and peak in mid to late August \citep{nault2011delaying}. While adults can be captured in fields using sticky cards, the larvae cause devastating damage by feeding on plant roots and onion bulbs \citep{nault2006onion, nault2006performance}. Feeding by \textit{D. antiqua} larvae early in the season can kill seedlings and damaged plants can become more susceptible to infection by pathogen and attack by subsequent generations \citep{eckenrode1986impact,nault2006performance}.  Feeding by \textit{D. antiqua} larvae, even if it does not completely kill the onion can cause sufficient damage to prevent sale of the produce.

Management of \textit{D. antiqua} has relied on prophylactic insecticide applications at planting \citep{nault2006performance}. A variety of organophosphate and carbamate insecticides have been used and discontinued \citep{nault2006performance}. Chlorpyrifos (Lorsban Advanced) is still commonly used as a drench treatment in combination with either cyromazine seed treatment (Trigard) \citep{nault2006performance,nault2006onion} or the seed treatment package containing thiamethoxam, spinosad and several fungicides (FarMore FI500); although some growers rely exclusively on FarMore FI500 to manage \textit{D. antiqua}. Cultural practices such as crop rotation, removal of damaged onions, and delayed planting are also effective \textit{D. antiqua} management tactics \citep{martinson1988dispersal,finch1985influence,nault2011delaying} but they are not typically adopted by large-scale commercial growers.

@@

Choice of the killing component in an attract and kill solution for use in an integrated pest management program is important. For working with \textit{D. antiqua} control, we elected to use spinosad as the toxicant. Spinosad is a naturally derived neurotoxic insecticide originally isolated from the soil-dwelling actinomycete \citep{Saccharopolyspora spinosa} \citep{racke2007reduced}. It has been approved for use on over 150 fruit and vegetable crops in the United States and is certified for use in organic production by the USDA National Organic Standards Board and Organic Materials Review Institute (OMRI) \citep{racke2007reduced,williams2003naturally}. In addition, spinosad has substantially lower non-target effects relative to conventional pesticides and a low residual toxicity \citep{williams2003naturally}. Because of its complementary nature to current conventional management of onion maggot and potential use in organic production systems, spinosad was included in the development of an attract and kill solution for \textit{D. antiqua}.

To develop an effective attract and kill solution to control \textit{D. antiqua}, we combined attractive white spheres developed in previous work \citep{willett2019} with a spinosad insecticide solution. We hypothesized that these spinosad-attractive spheres would kill \textit{D. antiqua} flies and significantly reduce maggot damage to onions. This study was conducted over the course of three field seasons in upstate New York.

@@

During the period these trials were conducted in upstate New York, \textit{Delia antiqua} had three generations (Figure \ref{fig:figure1}).  The first population of adults peaked in June, the second in late July, and the third in August.  More males than females were collected on sticky cards, but both males and females followed roughly the same population cycles.

@@

Comparisons of attract and kill strategies with insecticide and no insecticide controls suggest that spinosad containing spheres are controlling some damage by \textit{D. antiqua} larvae.  Spinosad spheres in conjunction with the attractive Delia Lure resulted in levels of damage almost as good as insecticide controls, but with variation that led to overlap with no insecticide negative controls.  In caged field trials, spinosad containing spheres reduced damage resulting from \textit{D. antiqua} feeding to below that of no insecticide controls.  Interestingly, pairing Delia Lure with sticky cards produced higher levels of damage.  This is likely because the kill part of the attract and kill sticky card solution was ineffective; Delia Lure likely attracted adult fly populations to plots but sticky cards did not cause sufficient mortality.  Augmenting spinosad spheres with ammonium carbonate and casein hydrolysate did not appreciably change control with this attract and kill method.

@@

Efficacy of this attract and kill strategy using spinosad containing spheres to control \textit{D. antiqua} could be enhanced through improving trap placement and density.  \textit{D. antiqua} adults are not uniformly distributed in onion fields, but tend to be found in higher numbers along field edges \citep{werling2006spatial}.  The numbers of the spinosad containing spheres and their placement along edges of commercial onion fields could be adjusted to have a greater impact on increasing fly mortality and reducing maggot damage.  A higher density of spheres in locations where maggot damage is a problem might have a more appreciable impact on reducing damage. Strategic placement of spinosad spheres in areas where populations are highest may further increase efficacy of this technique.  Additionally, increasing sphere density and temporally targeting the first \textit{D. antiqua} generations that cause the most damage may substantially improve efficacy of this technique and warrant further investigation.

>>>>

Onion maggot (\textit{Delia antiqua}) is a pest of onions worldwide. Current means of managing this pest rely heavily on prophylatic insecticide treatments at planting. These options may not be viable in organic production systems or situations where insecticide-resistant populations occur. Here we explore the efficacy of an attract and kill strategy for control of \textit{D. antiqua} evaluating the ability of attractive, spinosad containing spheres to kill adult \textit{D. antiqua} and reduce crop losses. Spinosad containing spheres were able to consistently kill \textit{D. antiqua} adults over the course of the field season (mortality range: between 49\% and 59\% on average). Pairing spinosad spheres with Delia Lure increased efficacy by 72\% compared with the attract and kill strategy alone. Performance of this attract and kill strategy also can reduce damage by \textit{D. antiqua}  larvae in the field, but it did not achieve a level of control comparable to the level provided by a conventional insecticide treatment. Implementation of this attract and kill strategy could be a valuable tool in situations where conventional pesticides are either not available or desired, where additional control techniques are needed, or to provide a season-long option for control of \textit{D. antiqua} populations.

@@

The lifecycle of \textit{Delia antiqua} plays a large role in its severity as an onion pest.  In the northern United States, three generations per year are common \citep{eckenrode1975population, hoepting2004insecticide}. In New York, \textit{D. antiqua} overwinters as a pupa and adults emerge in late April to early May and peak in late May to early June. The next generation begins to emerge in late June to early July and peaks in mid to late July, while the last generation adults begin to emerge in late July to early August and peak in mid to late August \citep{nault2011delaying}. While adults can be captured in fields using sticky cards, the larvae cause devastating damage by feeding on plant roots and onion bulbs \citep{nault2006onion, nault2006performance}. Feeding by \textit{D. antiqua} larvae early in the season can kill seedlings and damaged plants can become more susceptible to infection by pathogen and attack by subsequent generations \citep{eckenrode1986impact,nault2006performance}.  Feeding by \textit{D. antiqua} larvae, even if it does not completely kill the onion can cause sufficient damage to prevent sale of the bulb.

Management of \textit{D. antiqua} has relied on prophylactic insecticide applications at planting \citep{nault2006performance}. A variety of organophosphate and carbamate insecticides have been used and discontinued \citep{nault2006performance}. Chlorpyrifos (Lorsban Advanced) is still commonly used as a drench treatment in combination with either cyromazine seed treatment (Trigard) \citep{nault2006performance,nault2006onion} or the seed treatment package containing thiamethoxam, spinosad and several fungicides (FarMore FI500); although some growers rely exclusively on FarMore FI500 to manage \textit{D. antiqua}. Cultural practices such as crop rotation, removal of damaged cull and volunteer onions, and delayed planting are also effective \textit{D. antiqua} management tactics \citep{martinson1988dispersal,finch1985influence,nault2011delaying} but they are not typically adopted by large-scale commercial growers.

@@

Choice of the killing component in an attract and kill solution for use in an integrated pest management program is important. For working with \textit{D. antiqua} control, we elected to use spinosad as the toxicant. Spinosad is a naturally derived neurotoxic insecticide originally isolated from the soil-dwelling actinomycete \textit{Saccharopolyspora spinosa} \citep{racke2007reduced}. It has been approved for use on over 150 fruit and vegetable crops in the United States and is certified for use in organic production by the USDA National Organic Standards Board and Organic Materials Review Institute (OMRI) \citep{racke2007reduced,williams2003naturally}. In addition, spinosad has substantially lower non-target effects relative to conventional pesticides and a low residual toxicity \citep{williams2003naturally}. Because of its complementary nature to current conventional management of \textit{D. antiqua} and potential use in organic production systems, spinosad was included in the development of an attract and kill solution for \textit{D. antiqua}.

To develop an effective attract and kill solution to control \textit{D. antiqua}, we combined attractive white spheres developed in previous work \citep{willett2019} with a spinosad insecticide solution. We hypothesized that these spinosad-attractive spheres would kill \textit{D. antiqua} flies and significantly reduce maggot damage to onions in the field. This study was conducted over the course of three field seasons in upstate New York.

@@

During the period these trials were conducted in upstate New York, \textit{D. antiqua} had three generations (Figure \ref{fig:figure1}).  The first population of adults peaked in June, the second in late July, and the third in August.  More males than females were collected on sticky cards, but both males and females followed roughly the same population cycles.

@@

Comparisons of attract and kill strategies with insecticide and no insecticide controls suggest that spinosad containing spheres are reducing some damage by \textit{D. antiqua} larvae.  Spinosad spheres in conjunction with the attractive Delia Lure resulted in levels of onion damage almost as good as insecticide controls, but with variation that led to overlap with no insecticide negative controls.  In caged field trials, spinosad containing spheres reduced damage resulting from \textit{D. antiqua} feeding to below that of no insecticide controls.  Interestingly, pairing Delia Lure with sticky cards produced higher levels of damage.  This is likely because the kill part of the attract and kill sticky card solution was ineffective; Delia Lure likely attracted adult fly populations to plots but sticky cards did not cause sufficient mortality.  Augmenting spinosad spheres with ammonium carbonate and casein hydrolysate did not appreciably change control with this attract and kill method.

@@

Efficacy of this attract and kill strategy using spinosad containing spheres to control \textit{D. antiqua} could be enhanced through improving trap placement and density.  \textit{D. antiqua} adults are not uniformly distributed in onion fields, but tend to be found in higher numbers along field edges \citep{werling2006spatial}.  The numbers of the spinosad containing spheres and their placement along edges of commercial onion fields could be adjusted to have a greater impact on increasing fly mortality and reducing maggot damage.  A higher density of spheres in locations where maggot damage is a problem might have a more appreciable impact on reducing damage. Strategic placement of spinosad spheres in areas where populations are highest may further increase efficacy of this technique.  Additionally, increasing sphere density and temporally targeting the first \textit{D. antiqua} generation which causes the most damage, may substantially improve efficacy of this technique and warrant further investigation.
